Statically Typed String Sanitation Inside a Python

### 2014-08-21 sanitation-psp14/sanitation-psp14-short.tex
Added a TXT containing the submitted abstract, as well as a document for any errors we find before hearing about from reviewers. Also made   some minor changes.

\maketitle
\begin{abstract}
%Evidence suggests that programmers are reluctant to adopt new languages to
Web applications must ultimately command systems like web browsers and database engines using strings. Strings derived from improperly sanitized user input  can thus be a vector for command injection attacks. %Structured frameworks for constructing commands centralizes control over string sanitization, but verifying that the framework itself sanitizes user input correctly can still be tricky.
In this paper, we introduce \emph{regular string types}, which classify strings known statically to be in a specified regular language. These types come equipped with common operations like concatenation, substitution and coercion, so they  can be used to implement, in a conventional manner, the portions of a web application or application framework that must directly construct command strings. Simple type annotations at key interfaces can be used to statically verify that sanitization has been performed correctly without introducing redundant run-time checks. We specify this type system in a minimal typed lambda calculus, $\lambda_{RS}$.

To be practical, adopting a specialized type system like this should not require the adoption of a new programming language. Instead, we advocate for extensible type systems: new type system fragments like this should be implemented as libraries atop a mechanism that guarantees that they can be safely composed.  We support this by 1) specifying a translation from $\lambda_{RS}$ to a language fragment containing only standard strings and regular expressions, then, taking Python as a language with these constructs, 2) implementing the type system together with the translation as a library using \texttt{atlang}, an extensible static type system for Python being developed by the authors.% represents the most sophisticated application of this library to date.

% Security-oriented programming languages do not enjoy wide adoption. Conversely, library and framework-based approaches to security are pervasive. Unfortunately, libraries and frameworks are only effective solutions if they are implemented and used properly. In this paper, we propose that extensible programming languages provide a promising mechanism for increasing the amount of trust that can be place in libraries and frameworks. To demonstrate this approach, we present a conservative extension to the simply-typed lambda calculus for checking the correctness of input sanitation algorithms. We also demonstrate how this language's security guarantees are preserved under translation into an underlying language containing a regular expression library. We believe this approach -- complementing existing techniques with light-weight, composable type-based analyses -- constitutes a promising compromise between usability, potential for industry adoption, and theoretically grounded safety guarantees.
\end{abstract}

@@

In this paper, we present a static type system that tracks the regular language a string belongs to. For example, the output of \verb|sanitize| will be a string in the regular language described by the regular expression above (we identify regular languages by the notation $\mathcal{L}\{r\}$). By leveraging closure and decidability properties of regular languages, the type system tracks the language of a string through uses of a number of operations, including \emph{replacement} of substrings matching a given pattern. This makes it possible to implement sanitation functions like the one just described in a conventional manner. The result is a system where the fact that a string has been {correctly} sanitized is manifest in its type. Missing calls to sanitization functions are detected statically, and, importantly, so are \emph{incorrectly implemented sanitization functions} (i.e. these functions need not be trusted). These guarantees require run-time checks only when going from less precise to more precise types (e.g. at the edges of the system). %implementing and statically checking input sanitation techniques. Our solution suggests a more general approach to the integration of security concerns into programming language design. This approach is characterized by \emph{composable} type system extensions which \emph{complement} existing and well-understood solutions with compile-time checks.

@@

The \textsc{S-T-Concat} rule, which is defined in terms of these operations, relates the result of ``peeling off" the first character of a string to regular expression derivatives.

>>>>

\maketitle
\begin{abstract}
Web applications must ultimately command systems like web browsers and database engines using strings. Strings derived from improperly sanitized user input  can thus be a vector for command injection attacks. 
In this paper, we introduce \emph{regular string types}, which classify strings known statically to be in a specified regular language. These types come equipped with common operations like concatenation, substitution and coercion, so they  can be used to implement, in a conventional manner, the portions of a web application or application framework that must directly construct command strings. Simple type annotations at key interfaces can be used to statically verify that sanitization has been performed correctly without introducing redundant run-time checks. We specify this type system in a minimal typed lambda calculus, $\lambda_{RS}$.

To be practical, adopting a specialized type system like this should not require the adoption of a new programming language. Instead, we advocate for extensible type systems: new type system fragments like this should be implemented as libraries atop a mechanism that guarantees that they can be safely composed.  
We support this with two contributions. First, we specify a translation from $\lambda_{RS}$ to a language fragment containing only standard strings and regular expressions. Second, taking Python as a language with these constructs, we implement the type system together with the translation as a library using \texttt{atlang}, an extensible static type system for Python being developed by the authors.
\end{abstract}

@@

In this paper, we present a static type system that tracks the regular language a string belongs to. For example, the output of \verb|sanitize| will be a string in the regular language described by the regular expression above (we identify regular languages by the notation $\mathcal{L}\{r\}$). By leveraging closure and decidability properties of regular languages, the type system tracks the language of a string through uses of a number of operations, including \emph{replacement} of substrings matching a given pattern. This makes it possible to implement sanitation functions -- like the one just described -- in a conventional manner. The result is a system where the fact that a string has been {correctly} sanitized is manifest in its type. Missing calls to sanitization functions are detected statically, and, importantly, so are \emph{incorrectly implemented sanitization functions} (i.e. these functions need not be trusted). These guarantees require run-time checks only when going from less precise to more precise types (e.g. at the edges of the system). %implementing and statically checking input sanitation techniques. Our solution suggests a more general approach to the integration of security concerns into programming language design. This approach is characterized by \emph{composable} type system extensions which \emph{complement} existing and well-understood solutions with compile-time checks.

@@

The \textsc{S-T-Case-Concat} rule, which is defined in terms of these operations, relates the result of ``peeling off" the first character of a string to regular expression derivatives.
